---
title: "Practical 1"
---

```{r}
#| include: false
data(airquality)
airquality

```


# Airquality Data Exploration

### Missing Values

```{r}
#| echo: true
missingvalues<- airquality[!complete.cases(airquality), ]#gets rows that contains at least one NA.
missingvalues
```

## Summary statistics

### Temperature statistics
```{r}
#| echo: true 
Temperature<- airquality$Temp
Ozonelevel <- airquality$Ozone
tempmean <-mean(Temperature, na.rm=TRUE)
tempsd <- sd(Temperature, na.rm=TRUE)
tempmin<- min(Temperature, na.rm=TRUE)
tempmax<- max(Temperature, na.rm=TRUE)
TempStats<-c(tempmean, tempsd, tempmin,tempmax)
TempStats

```

### Ozone level statistics
```{r}
#| echo: true
Ozonemean <-mean(Ozonelevel, na.rm=TRUE)
Ozonesd <- sd(Ozonelevel, na.rm=TRUE)
Ozonemin<- min(Ozonelevel, na.rm=TRUE)
Ozonemax<- max(Ozonelevel, na.rm=TRUE)
OzoneStats<-c(Ozonemean, Ozonesd, Ozonemin,Ozonemax)
OzoneStats 
```


```{r}
#| include: false
data(cars)
```
# Cars Regression

## Regression analysis using Matrix method


```{r}
X<- cbind(1,cars$speed)# design matrix 
head(X)
Y<-cars$dist
head(Y)

```


The estimated regression coefficients are given by $\hat{\beta}$ 

$$\hat{\beta} = (X^TX)^{-1}X^TY$$

```{r}
X_transpose<-t(X)
f<- function(X,Y)
{bhat<-solve(X_transpose%*%X) %*% X_transpose%*%Y
bhat
Yhat<-X%*%bhat
Cmat <- solve(X_transpose%*%X)
Cmat
#k=p+1 where p=ncol of X
k <- 1+1
k
rss <- t(Y-X%*%bhat)%*%(Y-X%*%bhat)
rss
s2 <- as.numeric(rss)/(nrow(X)-k)
s2
#find diagonals of the C matrix 
c1 <- diag(Cmat)
c1
std.error <- sqrt(s2*c1)
std.error

return(list('coefficients'= bhat, 'std_error'=std.error))}
```

```{r}
#| echo: true
b<-f(X,Y)
b
```

## Regression analysis using lm()
Now using R lm() to get the coefficients.

```{r}
model1<-lm(dist~speed, data=cars)
summary(model1)
```
## Comparing coefficients 

```{r}
all.equal(as.numeric(b$coefficients), as.numeric(model1$coefficients))

```

## Conclusion 

The coefficient estimates produced by the two methods are the same.
